---
title: "p8106_midterm_zl2860"
author: "Zongchao Liu"
date: "4/1/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ModelMetrics)
library(caret)
```

# Import data

```{r, include=FALSE}
data = read_csv('./data/pokemon/pokemon.csv') %>% janitor::clean_names()
skimr::skim(data)
```

# Data at a glance

## correlation between basic pokemon traits and species strength

```{r}
library(corrplot)
num_predictor = data %>% 
  select("total", "hp", "attack", "defense", "sp_atk",
                           "sp_def", "speed", "generation","height_m","weight_kg", "catch_rate")

num_predictor = cor(num_predictor)
corrplot(num_predictor, method = "pie", type = "lower", 
         order = "FPC", tl.col = "grey", tl.srt = 45)
```


# visualize the catch rate

```{r,eval=FALSE}
data %>%
  ggplot() +
  #geom_histogram(aes(x = catch_rate,fill = generation))+
  geom_point(aes(x = total, y = catch_rate, color = generation)) +
  facet_grid(generation ~ .) +
  theme_bw() +
  labs(title = "Fig.2 Relationship Between Catch-rate and Species Strength",
       x = "Species Strength",
       y = "Catch Rate") +
  theme(plot.title = element_text(hjust = .5)) +
   ggsci:::scale_color_jama()

  
```

By linear regression, we find that the variable `total` is a sum of ``r colnames(num_predictor)[2:7]``. Hence, for further analysis we can substitute all of these variables with `total`.


# Reorganize the dataset for further analysis

```{r, include=FALSE}
set.seed(886)
data = read_csv('./data/pokemon/pokemon.csv') %>% 
  janitor::clean_names() %>%
  #select(-total) %>% #-hp,-attack,-defense,-sp_atk,-sp_def,-speed
  select(-number,-name,-type_2,-egg_group_2,-egg_group_1) %>%
  mutate(pr_male = ifelse(is.na(pr_male), -1, pr_male),
         generation = factor(generation))

index = sample(1:nrow(data),replace = F)
data = data[index,]

train = sample_frac(data,size = 0.8, replace = F)# training set 

validation = sample_frac(data,size = 0.2, replace = F) # validation set


train_x = model.matrix(catch_rate ~ . , train)[,-1]
train_y = train$catch_rate
val_x = model.matrix(catch_rate ~ . ,validation)[,-1]
val_y = validation$catch_rate

ctrl1  = trainControl(method = "cv")
```



# predict catching rate

## linear regression

```{r}

set.seed(886)

lm.fit = train(train_x, 
               train_y,
               method = "lm",
               trControl = ctrl1)
summary(lm.fit)

pred.lm = predict(lm.fit, val_x)

mse(val_y,pred.lm)
```

## ridge
 
```{r}
set.seed(886)
ridge.fit = train(x = train_x,
                  y = train_y,
                  method = "glmnet",
                  tuneGrid = expand.grid(alpha = 0,
                                         lambda = exp(seq(1,3,length = 200))),
                  preProc = c("center","scale"),
                  trControl = ctrl1)

ridge.fit$bestTune #lambda
ridge.plot = ggplot(ridge.fit,xTrans = function(x) log(x)) +
                  labs(title = "RMSE v.s. Regularization Parameter (Ridge)") +
  theme_bw() +
  theme(plot.title = element_text(hjust =.5, size = 10))
coef(ridge.fit$finalModel,ridge.fit$bestTune$lambda)
pred.ridge = predict(ridge.fit, val_x)
mse(val_y,pred.ridge)
```


## LASSO

```{r}
set.seed(886)
lasso.fit = train(train_x,
                  train_y,
                  method = "glmnet",
                  tuneGrid = expand.grid(alpha = 1,
                                         lambda = exp(seq(-1,1,length = 200))),
                  preProc = c("center","scale"),
                  trControl = ctrl1)


lasso.fit$bestTune

lasso.plot = ggplot(lasso.fit, xTrans = function(x) log(x)) +
                  labs(title = "RMSE v.s. Regularization (LASSO)") +
  theme_bw() +
  theme(plot.title = element_text(hjust =.5, size = 10))

coef = coef(lasso.fit$finalModel,lasso.fit$bestTune$lambda)
pred.lasso = predict(lasso.fit,val_x)
mse(val_y,pred.lasso)


```

## PCR

```{r}
set.seed(886)
pcr.fit = train(x = train_x, 
                 y = train_y,
                 method = "pcr",
                 tuneGrid = data.frame(ncomp = 1:ncol(train_x)),
                 tuneLength = length(train),
                 trControl = ctrl1,
                 preProc = c("center", "scale"))

pred.pcr = predict(pcr.fit,val_x)

pcr.fit$bestTune 

mse(val_y, pred.pcr)

pcr.plot = ggplot(pcr.fit, highlight = T) + theme_bw()+
  labs(title = "RMSE v.s. Number of Components (PCR)") +
  theme(plot.title = element_text(hjust =.5, size = 10))

library(patchwork)

lasso.plot + ridge.plot + pcr.plot
  
```

# mars
```{r}
mars_grid = expand.grid(degree = 1:4, 
                         nprune = 2:60)

set.seed(886)
mars.fit = train(train_x, train_y,
                  method = "earth",
                  tuneGrid = mars_grid,
                  trControl = ctrl1)

mars.plot = ggplot(mars.fit) + theme_bw() +
  labs(title = "RMSE v.s. Terms (MARS)") +
  theme(plot.title = element_text(hjust =.5))

mars.fit$bestTune
pred.mars = predict(mars.fit,val_x)
summary(mars.fit$finalModel) 
```


# GAM

```{r}
set.seed(886)
gam.fit = train(train_x, train_y,
                 method = "gam",
                 tuneGrid = data.frame(method = "GCV.Cp", select = c(TRUE,FALSE)),
                 trControl = ctrl1)

ggplot(gam.fit)

summary(gam.fit$finalModel)
gam.fit$finalModel
gam.fit$bestTune
pred.gam = predict(gam.fit,val_x)
```

# KNN
```{r}
set.seed(886)
knn.fit = train(catch_rate ~ .,
                data = train,
                tuneGrid = data.frame(k = 1:40),
                method = "knn",
                trControl = ctrl1)
summary(knn.fit)
knn.fit$bestTune
knn.plot = ggplot(knn.fit) + theme_bw() +
  labs(title = "RMSE v.s. Number of Neighbours (KNN)") +
  theme(plot.title = element_text(hjust =.5))

pred.knn = predict(knn.fit,validation)
mse(val_y,pred.knn)

mars.plot+knn.plot
```

# Comparison

```{r}
set.seed(886)
resamp = resamples(list(lasso = lasso.fit,
                        ridge = ridge.fit,
                        pcr = pcr.fit,
                        lm = lm.fit,
                        mars = mars.fit,
                        gam = gam.fit,
                        knn = knn.fit
                        ))

summary(resamp)

parallelplot(resamp,metric = "RMSE")
bwplot(resamp, metric = "RMSE",
       main = "Model Comparison Based on RMSE")
```



# discussion on the results

```{r}
result_cr = tibble(
  lm = pred.lm,
  lasso = pred.lasso,
  ridge = pred.ridge,
  pcr = pred.pcr,
  mars = pred.mars,
  gam = pred.gam,
  real = val_y,
  knn = pred.knn
) %>%
  pivot_longer(1:ncol(result_cr),
               names_to = "model",
               values_to = "catch_rate")

result_cr %>%
  ggplot(aes(x = catch_rate, fill = model)) +
  geom_density() +
  facet_grid(model ~ . ) +
  theme_bw() + 
  labs(title = "Density of the Catch Rate Predicted by the Models",
       x = "Catch rate",
       y = "Density") +
  theme(plot.title = element_text(hjust = .5)) +
  ggsci:::scale_fill_jama()
```

# stability of the prediction

```{r}
# calculate the catched probability
result_cp = tibble(
  lm = pred.lm,
  lasso = pred.lasso,
  ridge = pred.ridge,
  pcr = pred.pcr,
  mars = pred.mars,
  gam = pred.gam,
  real = val_y,
  knn = pred.knn
)

# define  scenarios
max_hp = 240
#current_hp = 50


# calculate the modified catch rate

cal_b = function(catch_rate,cur_hp,status = "slept"){
  
  max_hp = 240
  catch_correction = 1.5 #1.5 == super ball 2.5 advanced ball
  
  if(status == "frozen" | status == "slept"){
    s = 2}
  
  else{s = 1.5}
  return((3*max_hp - 2*cur_hp) * catch_rate * catch_correction * s / (3 * max_hp))
}



# determine caught or not
determination_b = function(b){
  return(ifelse(b > 255, 1, 0))
}


   
         
```


# simulation 

```{r}
test_model = function(){
  
  # define  scenarios
  max_hp = 240
  current_hp = seq(1,max_hp) # 1~ 240
  catch_correction = 1.5 # use great ball here
  
  prediction = NULL
  for (cur_hp in current_hp) {
    
    print(cur_hp)
    
    res = NULL
    res = data.frame(
      b.real = determination_b(cal_b(result_cp$real,cur_hp)),
         b.lm = determination_b(cal_b(result_cp$lm,cur_hp)),
         b.lasso = determination_b(cal_b(result_cp$lasso,cur_hp)),
         b.gam = determination_b(cal_b(result_cp$gam,cur_hp)),
         b.mars = determination_b(cal_b(result_cp$mars,cur_hp)),
         b.ridge = determination_b(cal_b(result_cp$ridge,cur_hp)),
         b.pcr = determination_b(cal_b(result_cp$pcr,cur_hp)),
         b.knn = determination_b(cal_b(result_cp$knn,cur_hp)))
    
    correct_rate = c(
                  lm = sum(res$b.real == res$b.lm)/144,
                  gam = sum(res$b.real == res$b.gam)/144,
                  ridge = sum(res$b.real == res$b.ridge)/144,
                  pcr = sum(res$b.real == res$b.pcr)/144,
                  lasso = sum(res$b.real == res$b.lasso)/144,
                  knn = sum(res$b.real == res$b.knn)/144,
                  mars = sum(res$b.real == res$y)/144)
    prediction = rbind(prediction,correct_rate)
    
  }
  return(data.frame(prediction))
}


df = test_model() #get the simulation results
df = df  %>%
    mutate(hp = c(1:nrow(df))) %>%
  pivot_longer(1:7,
               names_to = "model",
               values_to = "acc")
df %>% 
  ggplot(aes(x = hp, y = acc, color = model)) +
  geom_line() +
  theme_bw() +
  labs(title = "Prediction Accuracy v.s. Pokemons' Current HP") +
  ggsci::scale_color_jama() +
  theme(plot.title = element_text(hjust = .5))


```

